---
title: "Replication of Knaus 2024 Notebook"
author: "Aaron Sennhenn"
date: "2025-12-03"
output: html_document
---
## Load packages and data

First, load the relevant packages

```{r, warning = F, message = F}
if (!require("cobalt")) install.packages("cobalt", dependencies = TRUE); library(cobalt)
if (!require("grf")) install.packages("grf", dependencies = TRUE); library(grf)
if (!require("hdm")) install.packages("hdm", dependencies = TRUE); library(hdm)
if (!require("gridExtra")) install.packages("gridExtra", dependencies = TRUE); library(gridExtra)
if (!require("tidyverse")) install.packages("tidyverse", dependencies = TRUE); library(tidyverse)
if (!require("AER")) install.packages("AER", dependencies = TRUE); library(AER)
source("C:/Users/aaron/OneDrive/Desktop/Data Science Studium/Master Thesis/R/Ridge NNet/nn_functions.R")
```

Neural Net functions..

```{r}
source("C:/Users/aaron/OneDrive/Desktop/Data Science Studium/Master Thesis/R/Ridge NNet/nn_functions.R")
```



and the data

```{r}
data(pension) # Find variable description if you type ?pension in console
# Treatment
D = pension$p401
# Instrument
Z = pension$e401
# Outcome
Y = pension$net_tfa
# Controls
X = model.matrix(~ 0 + age + db + educ + fsize + hown + inc + male + marr + pira + twoearn, data = pension)
colnames(X) = c("Age","Benefit pension","Education","Family size","Home owner","Income","Male","Married","IRA","Two earners")
```

Define useful quantities and set the seed:

```{r}
# Helpful variables
N = length(Y)
ones = matrix(1,N,1)

# Choose unit for which causal/instrumental forest weights should be calculated
unit = 1

# Define number of folds for cross-fitting
nfolds = 2

set.seed(1234)
```

## One function to calculate them all

Proposition 1 in the paper shows that the outcome weights vector can be obtained in the general form
$$\boldsymbol{\omega'} = (\boldsymbol{\tilde{Z}'\tilde{D}})^{-1} \boldsymbol{\tilde{Z}'T}$$

where $\boldsymbol{\tilde{Z}}$, $\boldsymbol{\tilde{D}}$ and $\boldsymbol{T}$ are pseudo-instrument, pseudo-treatment and the transformation matrix, respectively. 

This motivates the definition of the `weight_maker(Dtilde,Ztilde,T_mat)` function taking the three generic inputs and producing the outcome weights vector:

```{r}
weight_maker = function(Dtilde,Ztilde,T_mat) {
  omega = (t(Ztilde) %*% Dtilde)^(-1) %*% t(Ztilde) %*% T_mat
  return(omega)
}
```

In the following, we just need to define the three components for the respective estimators and plug them into this generic function to obtain the outcome weights.


## Estimate all nuisance parameters and get outcome smoother matrices

Equation (8) in the paper shows which nuisance parameters are required for the estimators under consideration. Here, we estimate them all using honest random forest and 2-fold cross-fitting. Additionally, we extract the relevant smoother matrices $\boldsymbol{S}$, $\boldsymbol{S^d_0}$, $\boldsymbol{S^d_1}$, $\boldsymbol{S^z_0}$, and $\boldsymbol{S^z_1}$.

```{r}
# Get fold number
fold = sample(1:nfolds,N,replace=T)

#Hyperparameters
nn_hyps <- list(size=5, maxit=500, decay=0.01, linout=TRUE)

# Initialize nuisance parameters
Yhat = Yhatd0 = Yhatd1 = Yhatz0 = Yhatz1 = 
  Dhat = Dhatz0 = Dhatz1 = Zhat = rep(NA,N)

# Initialize smoother matrices
S = Sd0 = Sd1 = Sz0 = Sz1 = matrix(0,N,N)

for (i in 1:nfolds){
  #Calculate nuissance parameters and smoother matrices
  nn_Yhat = fit_RidgeNN(X[fold != i,], Y[fold != i], nn_hyps)
  Yhat[fold == i] = predict_RidgeNN(nn_Yhat, X[fold == i,])$predictions
  S[fold == i, fold != i] = as.matrix(get_nn_weights(nn_Yhat, X[fold == i,]))

  nn_Yhatd0 = fit_RidgeNN(X[fold != i & D==0,], Y[fold != i & D==0], nn_hyps)
  Yhatd0[fold == i] = predict_RidgeNN(nn_Yhatd0, X[fold == i,])$predictions
  Sd0[fold == i, fold != i & D==0] = as.matrix(get_nn_weights(nn_Yhatd0, X[fold == i,]))

  nn_Yhatd1 = fit_RidgeNN(X[fold != i & D==1,], Y[fold != i & D==1], nn_hyps)
  Yhatd1[fold == i] = predict_RidgeNN(nn_Yhatd1, X[fold == i,])$predictions
  Sd1[fold == i, fold != i & D==1] = as.matrix(get_nn_weights(nn_Yhatd1, X[fold == i,]))

  nn_Yhatz0 = fit_RidgeNN(X[fold != i & Z==0,], Y[fold != i & Z==0], nn_hyps)
  Yhatz0[fold == i] = predict_RidgeNN(nn_Yhatz0, X[fold == i,])$predictions
  Sz0[fold == i, fold != i & Z==0] = as.matrix(get_nn_weights(nn_Yhatz0, X[fold == i,]))

  nn_Yhatz1 = fit_RidgeNN(X[fold != i & Z==1,], Y[fold != i & Z==1], nn_hyps)
  Yhatz1[fold == i] = predict_RidgeNN(nn_Yhatz1, X[fold == i,])$predictions
  Sz1[fold == i, fold != i & Z==1] = as.matrix(get_nn_weights(nn_Yhatz1, X[fold == i,]))

  nn_Dhat = fit_RidgeNN(X[fold != i,], D[fold != i], nn_hyps)
  Dhat[fold == i] = predict_RidgeNN(nn_Dhat, X[fold == i,])$predictions

  nn_Dhatz0 = fit_RidgeNN(X[fold != i & Z==0,], D[fold != i & Z==0], nn_hyps)
  Dhatz0[fold == i] = predict_RidgeNN(nn_Dhatz0, X[fold == i,])$predictions

  nn_Dhatz1 = fit_RidgeNN(X[fold != i & Z==1,], D[fold != i & Z==1], nn_hyps)
  Dhatz1[fold == i] = predict_RidgeNN(nn_Dhatz1, X[fold == i,])$predictions

  nn_Zhat = fit_RidgeNN(X[fold != i,], Z[fold != i], nn_hyps)
  Zhat[fold == i] = predict_RidgeNN(nn_Zhat, X[fold == i,])$predictions
}

```


For those not familiar with smoothers, let's observe that they produce exactly the same predictions as the original `predict()` functions:
```{r}
cat("Smoother matrices replicate predicted nuisance parameter?",
 all(
  all.equal(as.numeric(S %*% Y), Yhat) == TRUE,
  all.equal(as.numeric(Sd0 %*% Y), Yhatd0) == TRUE,
  all.equal(as.numeric(Sd1 %*% Y), Yhatd1) == TRUE,
  all.equal(as.numeric(Sz0 %*% Y), Yhatz0) == TRUE,
  all.equal(as.numeric(Sz1 %*% Y), Yhatz1) == TRUE)
)
```

Note that the the neural net smoother matices are not affine:
```{r}
cat("Affine smoother matrices?",
all(
  all.equal(as.numeric(ones),as.numeric(rowSums(S))) == TRUE,
  all.equal(as.numeric(ones),as.numeric(rowSums(Sd0))) == TRUE,
  all.equal(as.numeric(ones),as.numeric(rowSums(Sd1))) == TRUE,
  all.equal(as.numeric(ones),as.numeric(rowSums(Sz0))) == TRUE,
  all.equal(as.numeric(ones),as.numeric(rowSums(Sz1))) == TRUE)
)
```

Finally, define the different IPW weights for later use:
```{r}
lambda1 = D / Dhat
lambda0 = (1-D) / (1-Dhat)
lambdaz1 = Z / Zhat
lambdaz0 = (1-Z) / (1-Zhat)
```

# Outcome weights of DML and GRF


Declare Nuissance Parameters:

```{r}
Uhat = Y - Yhat
Vhat = D - Dhat
Rhat = Z - Zhat
T_nn = diag(N) - S
```

```{r}
Uhat = Y - Yhat
Dtilde_nn = Vhat
Rhat = Z - Zhat
T_nn = diag(N) - S
```

## Partially linear IV regression

To obtain the standard partially linear IV regression, we first solve the canonical moment condition to see which value should be replicated:
```{r}
to_be_rep_plriv = mean(Rhat * Uhat) / mean(Rhat * Vhat)
to_be_rep_plriv
```

Compared to IF, we only need to pass a different pseudo-instrument to the `weight_maker` to replicate this number:
```{r}
# Modified pseudo-instrument
Ztilde_plriv = Rhat * ones
# Get outcome weights
omega_plriv = weight_maker(Dtilde_nn, Ztilde_plriv, T_nn) 
cat("ω'Y replicates moment based implentation?",all.equal(as.numeric( omega_plriv %*% Y ), to_be_rep_plriv))
```


# NOTEBOOK 2: 
```{r, warning = F, message = F}
if (!require("OutcomeWeights")) install.packages("OutcomeWeights", dependencies = TRUE); library(OutcomeWeights)
if (!require("hdm")) install.packages("hdm", dependencies = TRUE); library(hdm)
if (!require("grf")) install.packages("grf", dependencies = TRUE); library(grf)
if (!require("cobalt")) install.packages("cobalt", dependencies = TRUE); library(cobalt)
if (!require("tidyverse")) install.packages("tidyverse", dependencies = TRUE); library(tidyverse)
if (!require("viridis")) install.packages("viridis", dependencies = TRUE); library(viridis)
if (!require("gridExtra")) install.packages("gridExtra", dependencies = TRUE); library(gridExtra)
```

set.seed(1234)





